Text as Policy: \\ Measuring Policy Similarity through Bill Text Reuse

### 2016-09-28 manuscript/main.tex
more minor revisions

\subsection{Data}
We apply the SW algorithm to a database of US state bills, collected by
\citet{burgess2016legislative} and the Sunlight
Foundation\footnote{\url{http://sunlightfoundation.com/}}. This data base
contains approximately 500,000 bills from 2008 to 2015\footnote{Since
    computation is very demanding for this problem, the results reported in this
    version of the paper are based on a random sample of about half of the
    bills. We are in the process of completing the analyses for the whole
dataset.}. Not all bills from all
states are available for the whole time period. Figure \ref{fig:bill_desc}
displays the year ranges and number of bills in the bill data base. This
collection of bills is based on all bills that are available through the
Sunlight Foundation's \url{openstates.org} API. \url{openstates.org} is a website maintained by the Sunlight Foundation, in order to increase transparency in state politics. The Sunlight Foundation uses web scrapers to access all bills that are available on the websites of legislatures in all US states. This includes enacted legislation as well as bills that are still in the legislative process, or where not enacted. %The database contains the bill text as well as additional information on the legislative process for the bill. We have the first and the last version of the bill text, the sponsor of the bill and all legislative action that was taken on the bill. %The available meta-data for the bills includes a time-stamp for introduction and approval of the bill, the name and party affiliation of the sponsor(s), the state and the bill id.

\begin{figure}[ht!]
    \centering
    \includegraphics[width=0.65\textwidth]{figures/year_count_by_state.png}
    \caption{Description of the bill database. The horizontal bars display the time range for which bills are available for each state. The color of the bar indicates how many bills are available for this state.}
    \label{fig:bill_desc}
\end{figure}

@@

The SW algorithm is computationally demanding and an exhaustive comparison of
all possible bill-pairs is not feasible (with $500,000$ bills this would amount
to about $1.24 * 10^11$ bill pairs). \citet{wilkerson2015tracing} encountered the same problem and excluded bills that don't have at least five matching 10-grams. Since this implies considerable exact character matches, we see this as a too restrictive for our application. Because the text we are using originates in 50 different legislatures, exact character matches might obfuscate valuable text reuse just because of small formatting changes, changes of state names, etc. 
Instead we choose to rely on a procedure that is based on finding bills with
similar language. We store all available bills in an Elastic Search database.
Elastic search is an open source web search engine which is designed to find
documents that match a search query \citep{gormley2015elasticsearch}. We utilize
the `more like this' query, which is designed to find documents that are similar
to a given document. It works in the following way: From the focus document, the
$k$ $n$-grams with the highest tf-idf scores are selected and transformed into a
search vector that is representative of the query document. Then, for each
document in the collection, a variation of the cosine similarity is calculated
and the $m$ bills with the highest scores are selected for further analysis. For
this analysis we set $k$ to 25, $n$ to 5 and $m$ to 1000\footnote{For
details on the exact algorithm we refer the reader to the elastic search
documentation
(\url{https://www.elastic.co/guide/en/elasticsearch/reference/current/query-dsl-mlt-query.html} as well as the lucene engine
documentation
(\url{https://lucene.apache.org/core/4_9_0/core/org/apache/lucene/search/similarities/TFIDFSimilarity.html}).
In order to account for the possibility of selection bias being introduced to
our analyses through the cap of 1000 similar bills, we analyzed the results with
different numbers of pre-selected bills. We find that this parameter, when set
to a sufficiently high value, does not lead to such bias because the set of
potentially relevant bills is always much smaller than 1000.}.

@@

In the second approach we describe, which we have implemented in the analyses we present below, we do not exclude boiler plate according to a classification dichotomy.
Rather, we down-weight alignments that consist of language that is
very common to all alignments. The underlying logic is that language that
appears in a large number of other alignments is probably not related to
specific policies. For example, alignments consisting of many stop words such as
`the' or `a' or of common procedural language like `legislature` or `section'
are most likely part of phrases that are either very frequent in common language
or very frequent in the legislative context. 
In order to quantify the similarity to common alignments, we calculate the
cosine similarity of the bag-of-words vector of each alignment with a random
sample of 1000 alignments\footnote{We rely on a random sample for computational
reasons. Calculating the complete similarity matrix is computationally
infeasible and unnecessary.}. The adjusted score for two bills $X$ and $Y$ is
then:

@@

Figure \ref{fig:ncsl_prec_rec} displays the performance of the alignment score
in predicting if two bills are in the same NCSL table. We present
precision---the proportion of bill pairs with a given alignment score that
exceed the threshold on the $x$-axis that are in the same NCSL table, and
recall---the proportion of same-NCSL-table pairs that exceed a given alignment
score threshold. We see very high precision when setting the threshold to
relatively high scores ($>100$). Recall is low at thresholds that produce
reasonable precision (above 100), but that is expected for our case, as
(1) the introduction of similar policies in bills in tow different states is a
rare event, and relatedly, (2) the appearance of two bills in the same NCSL
table is a relatively rare event (8\% in the current analysis); classifier
recall is notoriously low in rare event data \citep{weiss2000learning}---even
commonly approaching zero, as in our case \citep{weiss2004mining}. Further more,
using a high precision - low recall classifier in our scenario where the total
number of bill pairs is extremely large, produces a large number of interesting
bill pairings in absolute terms.

@@

To evaluate whether text re-use corresponds to the transfer of policy, we test whether the presence of a diffusion network tie between two states is a predictor of text reuse. We use the policy diffusion networks inferred in \citet{desmarais2015}. The diffusion networks were inferred using policy adoption sequences, and the network inference algorithm developed by \citet{gomez2010inferring}. A tie from state $i$ to state $j$ in the diffusion network indicates that state $j$ has frequently emulated state $i$'s policies in the preceding thirty-five years. To calculate an aggregate alignment score for each state-pair we calculate the sum of the natural logs of the alignment scores associated with each pair of bills across the two states. We use the log since a handful of large alignment scores lead to extreme outliers on the original scales. The ``Diffusion Ties'' variable indicates the presence of a diffusion edge between states in the 2008 diffusion network, as measured by \citet{desmarais2015}. The diffusion network in 2008 is inferred using policy adoptions in the 35 years preceding (and excluding) 2008.  There is one observation in the analysis for each of the 1,225 unique state-pairs. Since this is dyadic data, we use a matrix permutation method, quadratic assignment procedure, to calculate $p$-values \citep{krackhardt1988}. As a robustness check, we run the model with both the identity and log link.\footnote{The $p$-values were calculated using 5,000 random matrix permutations.} We also control for a dyadic variable (Coverage) equal to the product of the number of years for which we have data for each state in the dyad.

% latex table generated in R 3.2.3 by xtable 1.8-2 package
% Wed Sep 21 20:12:14 2016
\begin{table}[ht]
\centering
\begin{tabular}{rrrrr}
\hline
& \multicolumn{2}{c}{Identity Link} & \multicolumn{2}{c}{Log Link} \\
  \hline
    & Coefficient & p-value & Coefficient & p-value \\
    Intercept & -3.297 & 0.997 & 9.014 & 0.000 \\ 
    Diffusion Tie & 0.238 & 0.042 & 0.239 & 0.050 \\ 
    Coverage & 1.074 & 0.000 & 1.238 & 0.000 \\ 
\hline
\end{tabular}
\caption{Predicting aggregate cross-state alignment scores with diffusion ties.
Coefficients calculated with OLS regression and normalized with standard
deviation of cross state alignment scores. $p$-values based on 5,000 QAP permuations.}
\label{tab:qap.diffusion}
\end{table}

Results of the dyadic regression are presented in Table \ref{tab:qap.diffusion}. In both specifications there is a positive relationship between the number of diffusion ties and the number of alignments. The relationship is statistically significant at the 0.05 level (one tailed) in each analysis. Furthermore, the magnitudes of the relationships are substantively significant. With the identity link, the presence of a diffusion tie leads to a 0.23 standard deviation increase in the expected aggregate alignment score. Based on the log link, the addition of a diffusion tie corresponds to a 30\% increase in the expected aggregated alignment score.\footnote{Calculated as $100\times \left[ \exp(0.266)-1\right] = 30.47$}. These results offer further evidence of the validity of quantifying text re-use as a measure of policy similarity, as the aggregate volume of text re-use between states is positively associated with previously-identified diffusion ties between states.

@@

In the following sections we present several analyses to assess this correlation. The ideal point of the bill is derived from the ideal point of the primary bill sponsor. For bills with several primary sponsors we use the average of those sponsors' ideal points to obtain a single measure. Figure \ref{fig:ideology_plot} displays the distribution of ideological distance and log-alignment score for the roughly 50 million dyads in the dataset. Each dot is a hexagonal tile, the color of the tile represents the number of dyads in that area. Two major observations can be made in this plot. First, the vast majority of alignment scores are close to zero. Second, the triangular shape of the distribution indicates the high precision - low recall character of the alignment score discussed above: few bills exhibit meaningful policy overlap with other bills, but in the region of lower ideological distance we observe a greater occurrence of positive outliers, which arise from significant policy overlap.

>>>>

\subsection{Data}
We apply the SW algorithm to a database of US state bills, collected by
\citet{burgess2016legislative} and the Sunlight
Foundation.\footnote{\url{http://sunlightfoundation.com/}} This data base
contains approximately 500,000 bills from 2008 to 2015\footnote{Since
    computation is very demanding for this problem, the results reported in this
    version of the paper are based on a random sample of about half of the
    bills. We are in the process of completing the analyses for the whole
dataset.}. Not all bills from all
states are available for the whole time period. Figure \ref{fig:bill_desc}
displays the year ranges and number of bills in the bill data base. This
collection of bills is based on all bills that are available through the
Sunlight Foundation's \url{openstates.org} API. \url{openstates.org} is a website maintained by the Sunlight Foundation, in order to increase transparency in state politics. The Sunlight Foundation uses web scrapers to access all bills that are available on the websites of legislatures in all US states. This includes enacted legislation as well as bills that are still in the legislative process, or where not enacted. %The database contains the bill text as well as additional information on the legislative process for the bill. We have the first and the last version of the bill text, the sponsor of the bill and all legislative action that was taken on the bill. %The available meta-data for the bills includes a time-stamp for introduction and approval of the bill, the name and party affiliation of the sponsor(s), the state and the bill id.

\begin{figure}[ht!]
    \centering
    \includegraphics[width=0.8\textwidth]{figures/year_count_by_state.png}
    \caption{Description of the bill database. The horizontal bars display the time range for which bills are available for each state. The color of the bar indicates how many bills are available for this state.}
    \label{fig:bill_desc}
\end{figure}

@@

The SW algorithm is computationally demanding and an exhaustive comparison of
all possible bill-pairs is not feasible (with $500,000$ bills this would amount
to about $1.24 * 10^11$ bill pairs). \citet{wilkerson2015tracing} encountered the same problem and excluded bills that don't have at least five matching 10-grams. Since this implies considerable exact character matches, we see this as a too restrictive for our application. Because the text we are using originates in 50 different legislatures, exact character matches might obfuscate valuable text reuse just because of small formatting changes, changes of state names, etc. 
Instead we choose to rely on a procedure that is based on finding bills with
similar language. We store all available bills in an Elastic Search database.
Elastic search is an open source web search engine which is designed to find
documents that match a search query \citep{gormley2015elasticsearch}. We utilize
the `more like this' query, which is designed to find documents that are similar
to a given document. It works in the following way: From the focus document, the
$k$ $n$-grams with the highest tf-idf scores are selected and transformed into a
search vector that is representative of the query document. Then, for each
document in the collection, a variation of the cosine similarity is calculated
and the $m$ bills with the highest scores are selected for further analysis. For
this analysis we set $k$ to 25, $n$ to 5 and $m$ to 1000.\footnote{For
details on the exact algorithm we refer the reader to the elastic search
documentation
(\url{https://www.elastic.co/guide/en/elasticsearch/reference/current/query-dsl-mlt-query.html} as well as the lucene engine
documentation
(\url{https://lucene.apache.org/core/4_9_0/core/org/apache/lucene/search/similarities/TFIDFSimilarity.html}).
In order to account for the possibility of selection bias being introduced to
our analyses through the cap of 1000 similar bills, we analyzed the results with
different numbers of pre-selected bills. We find that this parameter, when set
to a sufficiently high value, does not lead to such bias because the set of
potentially relevant bills is always much smaller than 1000.}

@@

In the second approach we describe, which we have implemented in the analyses we present below, we do not exclude boiler plate according to a classification dichotomy.
Rather, we down-weight alignments that consist of language that is
very common to all alignments. The underlying logic is that language that
appears in a large number of other alignments is probably not related to
specific policies. For example, alignments consisting of many stop words such as
`the' or `a' or of common procedural language like `legislature` or `section'
are most likely part of phrases that are either very frequent in common language
or very frequent in the legislative context. 
In order to quantify the similarity to common alignments, we calculate the
cosine similarity of the bag-of-words vector of each alignment with a random
sample of 1000 alignments.\footnote{We rely on a random sample for computational
reasons. Calculating the complete similarity matrix is computationally
infeasible and unnecessary.} The adjusted score for two bills $X$ and $Y$ is
then:

@@

Figure \ref{fig:ncsl_prec_rec} displays the performance of the alignment score (left panel)
in predicting if two bills are in the same NCSL table compared to the performance of a simple cosine similarity measure (right panel). We present
precision---the proportion of bill pairs with a given alignment score that
exceed the threshold on the $x$-axis that are in the same NCSL table, and
recall---the proportion of same-NCSL-table pairs that exceed a given alignment
score threshold. We see very high precision when setting the threshold to
relatively high scores ($>100$). Recall is low at thresholds that produce
reasonable precision (above 100), but that is expected for our case, as
(1) the introduction of similar policies in bills in tow different states is a
rare event, and relatedly, (2) the appearance of two bills in the same NCSL
table is a relatively rare event (8\% in the current analysis); classifier
recall is notoriously low in rare event data \citep{weiss2000learning}---even
commonly approaching zero, as in our case \citep{weiss2004mining}. Further more,
using a high precision - low recall classifier in our scenario where the total
number of bill pairs is extremely large, produces a large number of interesting
bill pairings in absolute terms.

@@

To evaluate whether text re-use corresponds to the transfer of policy, we test whether the presence of a diffusion network tie between two states is a predictor of text reuse. We use the policy diffusion networks inferred in \citet{desmarais2015}. The diffusion networks were inferred using policy adoption sequences, and the network inference algorithm developed by \citet{gomez2010inferring}. A tie from state $i$ to state $j$ in the diffusion network indicates that state $j$ has frequently emulated state $i$'s policies in the preceding thirty-five years. To calculate an aggregate alignment score for each state-pair we calculate the sum of the alignment scores associated with each pair of bills across the two states. The ``Diffusion Ties'' variable indicates the presence of a diffusion edge between states in the 2008 diffusion network, as measured by \citet{desmarais2015}. The diffusion network in 2008 is inferred using policy adoptions in the 35 years preceding (and excluding) 2008.  There is one observation in the analysis for each of the 1,225 unique state-pairs. Since this is dyadic data, we use a matrix permutation method, quadratic assignment procedure, to calculate $p$-values \citep{krackhardt1988}. As a robustness check, we run the model with both the identity and log link.\footnote{The $p$-values were calculated using 5,000 random matrix permutations.} We also control for a dyadic variable (Coverage) equal to the product of the number of years for which we have data for each state in the dyad.

% latex table generated in R 3.2.3 by xtable 1.8-2 package
% Wed Sep 21 20:12:14 2016
\begin{table}[ht]
\centering
\begin{tabular}{rrrrr}
\hline
& \multicolumn{2}{c}{Identity Link} & \multicolumn{2}{c}{Log Link} \\
  \hline
    & Coefficient & p-value & Coefficient & p-value \\
    Intercept & -3.297 & 0.997 & 9.014 & 0.000 \\ 
    Diffusion Tie & 0.238 & 0.042 & 0.239 & 0.050 \\ 
    Coverage & 1.074 & 0.000 & 1.238 & 0.000 \\ 
\hline
\end{tabular}
\caption{Predicting aggregate cross-state alignment scores with diffusion ties.
Coefficients calculated with OLS regression and normalized with standard
deviation of cross state alignment scores. One-tailed $p$-values based on 5,000 QAP permuations.}
\label{tab:qap.diffusion}
\end{table}

Results of the dyadic regression are presented in Table \ref{tab:qap.diffusion}. In both specifications there is a positive relationship between the number of diffusion ties and the number of alignments. The relationship is statistically significant at the 0.05 level (one tailed) in each analysis. Furthermore, the magnitudes of the relationships are substantively significant. With the identity link, the presence of a diffusion tie leads to a 0.23 standard deviation increase in the expected aggregate alignment score. Based on the log link, the addition of a diffusion tie corresponds to a 30\% increase in the expected aggregated alignment score.\footnote{Calculated as $100\times \left[ \exp(0.266)-1\right] = 30.47$.} These results offer further evidence of the validity of quantifying text re-use as a measure of policy similarity, as the aggregate volume of text re-use between states is positively associated with previously-identified diffusion ties between states.

@@

In the following sections we present several analyses to assess this correlation. The ideal point of the bill is derived from the ideal point of the primary bill sponsor. For bills with several primary sponsors we use the average of those sponsors' ideal points to obtain a single measure. Figure \ref{fig:ideology_plot} displays the distribution of ideological distance and log-alignment score for the roughly 50 million dyads in the dataset. The left panel displays the alignment scores on the natural scale. In order to get a better view of the distribution for small alignments, the right panel on the logarithmic scale. Each dot is a hexagonal tile, the shading of the tile represents the density in the area of the tile. Two major observations can be made in this plot. First, the vast majority of alignment scores are close to zero. Second, the triangular shape of the distribution indicates the high precision - low recall character of the alignment score discussed above: few bills exhibit meaningful policy overlap with other bills, but in the region of lower ideological distance we observe a greater occurrence of positive outliers, which arise from significant policy overlap.
